### 2025-06-08 thesis/thesis.tex
Write a very short intro to transformers

@@

% Conceptual Design
\subsection{System Design}

@@

% Implementation
\section{Implementation}

>>>>

\subsubsection{Neural Networks}

% already wrote about NNs and their foundations

\subsubsection{Sequence to Sequence Architecture}

\subsubsection{Transformer Architecture}

The transformer architecture advanced the sequence processing capabilities of neural networks and enabled parallelized inference
which was a critical limitation in efficiency of RNNs and LSTMs. Namely the introduction of attention (sometimes referenced as self-
attention) allow transformers led to this leap in capability \citep{Attention}. The core architectural characteristics of transformers
include:

\begin{enumerate}
	\item \textbf{Input Embedding} — Techniques like \textsc{Word2Vec} allow the semantic representation of words and tokens in 
		embeddings (ie. vectors in a N-dimensional space).
	\item \textbf{Positional Encoding} — Allows transformers to keep track of each tokens position in a sequence by adding the
	 	output of an offset function onto the original embedding.
	\item \textbf{Self-Attention} — Essentially weighting the relationship between tokens and the semantic importance of tokens.
	    Attention allows for both parallelized encoding, as well as semantic aware decoding, by calculating encoder-decoder attention 
	    prior to decoding.              
\end{enumerate}

Due to these architectural properties, transformers are better positioned to capture long-range dependencies in input sequences,
and preserving them into the decoding phase – rendering them as the state-of-the-art model architecture for NL2SQL.

@@

% Conceptual Design

\subsection{System Design}

@@

% Implementation

\section{Implementation}
